
\section{Elementary Definitions}

\subsection{Independent sets}

We will begin with the first possible way of defining what a \textit{matroid} is. This way is arguably the simplest one because the properties are intuitive and not hard to visualize. When speaking about matroids, we will always deal with finite sets, and the way we obtain a matroid from a finite set is to select some special selection of its subsets.  These special sets correspond to the \textit{independent} sets, and should obey some distinctive properties. This idea is precisely what the following definition is about.

\begin{defn}
    Let $E$ be a finite set, possibly empty, and $\mathcal{I}$ a collection of subsets of $E$ (i.e. some subset of the power set $2^E$ of E). We call the ordered pair $M = (E, \mathcal{I})$ a matroid if the following three properties are satisfied

    \begin{enumerate}
        \item[(I1)] We have $\emptyset \in \mathcal{I}$.
        
        \item[(I2)] If $I \in \mathcal{I}$ and $J \subseteq I$, then $J \in \mathcal{I}$.
        
        \item[(I3)] If $J, I \in \mathcal{I}$ and $|J| < |I|$, then there exists $e \in I - J$ so that $J \cup e \in \mathcal{I}$.
    \end{enumerate}

    We call elements of $\mathcal{I}$ \textbf{independent sets}. 
\end{defn}

There are some alternatives to the third property that are equivalent, for example we could alternatively write

\textit{3.* If $I, J \in \mathcal{I} $ and $|J| = |I| + 1$, then there exists $e \in J - I$ such that $I \cup e \in \mathcal{I}$.}

or

\textit{3.** If $X \subseteq E$ and $I_1, I_2$ are maximal members of $\{ I \in \mathcal{I} | I \subseteq X \}$, then $|I_1| = |I_2|$.}

In  figure \ref{first1234} you find a representation of a matroid by writing all subsets of the ground set $\{1,2,3,4\}$ and coloring them accordingly. The independent sets are colored in orange. It is clear that the empty set is independent and that all subsets of independent sets are independent. The dependent sets, colored in cyan, are the sets that are not independent.


\begin{figure}[h]

\begin{center}
\begin{tikzpicture}

\matrix (a) [matrix of math nodes, column sep=0.6cm, row sep=0.6cm,]{
 & & &\textcolor{cyan}{
1234} & & & &\\
 \textcolor{cyan}{
123}& &\textcolor{cyan}{
124} & &\textcolor{cyan}{
134} &  & \textcolor{cyan}{
234}  \\
\textcolor{orange}{12} & \textcolor{cyan}{13} & \textcolor{cyan}{14} & & \textcolor{orange}{23} & \textcolor{cyan}{
24} & \textcolor{cyan}{
34} \\
\textcolor{orange}{1}& &\textcolor{orange}{2} & & \textcolor{orange}{3}& & \textcolor{cyan}{4} \\
& & & \textcolor{orange}{\emptyset} &  & & \\
&&&&&& \\};

\foreach \i/\j in {1-4/2-1, 1-4/2-3, 1-4/2-5, 1-4/2-7, 2-1/3-1, 2-1/3-2, 2-1/3-5, 2-3/3-1, 2-3/3-3, 2-3/3-6, 2-5/3-2, 2-5/3-3, 2-5/3-7, 2-7/3-5, 2-7/3-6, 2-7/3-7, 3-1/4-1, 3-1/4-3, 3-2/4-1, 3-2/4-5, 3-3/4-1, 3-3/4-7, 3-5/4-3, 3-5/4-5, 3-6/4-3, 3-6/4-7, 3-7/4-7, 3-7/4-5, 4-1/5-4, 4-3/5-4, 4-5/5-4, 4-7/5-4}
\draw[double, line width = 0.005mm, color = brown] (a-\i) -- (a-\j);

\node[draw] at (0, -2.5){\small \textcolor{orange}{Independent set}, \textcolor{cyan}{Dependent set}};

\end{tikzpicture} % why does it say this??? Did you forget a semicolon?
\end{center}
\caption{Diagram showing independent and dependent sets of a 4-element matroid.}
\label{first1234}
\end{figure}


The definition of a matroid is designed to extract out the property that makes a subset of elements 'independent'. This leads us to our first examples, namely the so-called \textit{representable matroids} arising from linear algebra. We would that a subset of vectors is \textit{independent} in a matroid, if and only if it is are \textit{linearly independent} in the usual sense. 
%This means, in particular, that the subsets have the same properties.



 Let $A \in \mat_{m \times n}(\mathbb{F})$ where $\mathbb{F}$ is a field. In the article we will not just be interested with $\mathbb{F} = \mathbb{C} \; \mathrm{or}\; \mathbb{F} = \mathbb{R} $ but also in finite fields 

 
 \footnote{In particular, we are interested in $\mathbb{Z} / p\mathbb{Z}$. By that we mean $\mathbb{Z}$ modulo $p\mathbb{Z}$ so the set of reminders when dividing by a prime number $p$. We will denote this field by $\mathbb{F}_p$}.

\begin{exmp}
    

 
 We will pick a concrete example of $A \in  \mat_{3 \times 4}(\mathbb{R})$ to illustrate that the set of columns of a matrix has a natural matroid structure. Suppose

$$A = \begin{pmatrix}

2 & 0 & 2 & 0 \\
1 & -1 & 0 & 0 \\
0 & 3 & 3 & 0


\end{pmatrix}.$$

We would like to consider \textit{the set of labels of columns of the matrix $A$} and form a matroid on it. This means we start with $E = \{1,2,3,4\}$ so a finite set where the number $1$ corresponds to the column $\begin{pmatrix} 2 & 1 & 0 \end{pmatrix} ^ T$, the number 2 corresponds to $\begin{pmatrix}  0 & -1 & 3  \end{pmatrix} ^ T$ and so forth. We now declare that a subset of $E$ is called independent if and only if the corresponding set of column vectors is linearly independent as a set of vectors in $\mathbb{R}^3$. We can explicitly check what this means in our example. The sets $\{1\}$, $\{2\}$, $\{3\}$ are all independent because they correspond to non-zero vectors. For the two-element subsets we see that $\{1,2\}$, $\{1,3\}$ and $\{2,3\}$ are all independent, while any two element subsets containing the last column are not. Finally, the 3-element subset $\{1,2,3\}$ is not independent because as vectors, the first and the second column sum up to the third. So to conclude, the collection of all independent sets is
$$\mathcal{I} = \{\emptyset, \{1\}, \{2\}, \{3\}, \{1,2\}, \{1,3\}, \{2,3\} \},$$

which indeed satisfies the properties of collection of independent sets of a matroid. This is not a coincidence, we will prove that a collection of subsets formed from a matrix in the above way is always a matroid.

\end{exmp}


\begin{theorem}

    Let $A \in \mathrm{Mat} _{m \times n}(\mathbb{F})$ and $E = \{1, 2, \cdots, n\}$ be a finite set of $n$ elements where the element $i$ corresponds to the $i$-th column of the matrix $A$. We call a subset $I \subseteq E$ independent if and only if the column vectors that the members of $I$ correspond to form a linearly independent set as members of $\mathbb{F}^n$, and denote the collection of all independent subsets as $\mathcal{I}$. Then $M = (M, \mathcal{I})$ is a matroid, and we denote it by $M[A]$.
    
\end{theorem}

\begin{proof}
    We need to check that the collection $\mathcal{I}$ satisfies the three properties for the collection of independent sets given in the definition. 

\begin{enumerate}

    \item First, $\emptyset$ is trivially in $\mathcal{I}$.
    
    \item The second property is also satisfied because if $J \subseteq I$ and $I \in \mathcal{I}$ this means that the vectors corresponding to $I$ form a linearly independent set. In particular, let $v_1, v_2, \cdots v_j, v_{j+1}, \cdots v_{i}$ be the column vectors corresponding to $I$, and that first $j$ are also in $J$, that means $|I| = i$, $|J| = j$ and $j \leq i$. If for some linear combination we have $a_1v_1 + \cdots + a_jv_j = 0$, where $a_i \in \mathbb{F}$ then it is also true that $a_1v_1 + \cdots + a_jv_j + 0 \cdot a_{j+1} + \cdots + 0 \cdot a_i = 0$. Since the vectors of $I$ are linearly independent it now follows that $a_1 = a_2 = \cdots a_j = 0$ as well. So the vectors corresponding to the elements of $J$ are also linearly independent, which means by definition $J \in \mathcal{I}$.

    \item Finally, we have to check third property. We assume that $J, I \in \mathcal{I}$ and $|J| < |I|$. We denote by $V_J$ and $V_I$ the linear subspaces of $\mathbb{F}^n$ spanned by vectors corresponding to $J$ and $I$ respectively. Because the vectors corresponding to $J$ and $I$ are linearly independent, they also form a basis for $V_J$ and $V_I$ respectively. For any $e \in I - J$ we denote by $v(e) \in \mathbb{F}^n$ the column vector of $A$ corresponding to $e$. Suppose some $e \in I - J$ has the property that adding it to $J$ does not change the dimension of the subspace spanned by J, i.e. $\dim (V_J \cup v(e)) = \dim (V_J) = |J|$. Then $v(e)$ is already in $V(J)$ because the vectors corresponding to $J$ are linearly independent and if the dimension does not increase, then $v(e)$ can be expressed as a linear combination of vectors corresponding to $J$. However, this cannot hold true for \textit{every} $e \in I - J$. Suppose for contradiction, that for every $e \in I - J$ it holds that the vector $v(e) $ is already in $ V_J$. Because for all elements in the intersection,  $i \in I \cap J$, we have $v(i) \in V_J$ by definition, we could then infer $V_I \subseteq V_J$. But this implies that 

    $$|I| = \dim(V_I) = \dim(V_J) = |J| < |I|$$
     which is a contradiction. So there is at least one $e \in I - J$ so that $\dim(V_J \cup v(e)) = \dim(V_J) + 1$, which means the vectors corresponding to $J \cup e$ form a linearly independent subset. Finally, this means that $J \cup e \in \mathcal{I}$ which proves the third property.

\end{enumerate}
\end{proof}

In order to talk about any classification of matroids we have to say when the two matroids are equivalent - representing the same structure of independent sets. Intuitively, it is nothing deep, the definition will just rephrase that the two are \textit{equal} if it is possible to relabel the elements of the first matroid to the elements of the second without changing the independent sets.

\begin{defn}
    We call two matroids $M = (E, \mathcal{I})$ and $N = (F, \mathcal{J})$ isomorphic and denote it by $M \sim N$ if there exists a bijection $f: E \to F$ so that a subset $K \subseteq F$ is independent if and only if $K = f(L)$ for some independent set $L \in \mathcal{I}$.
\end{defn}


\begin{defn}
    We call a matroid $M$ representable, if $M$ is isomorphic to a matrix matroid $N[A]$ for some $A \in \mat_{m \times n}(\mathbb{F})$ over some field $\mathbb{F}$ and we call it $\mathbb{F}$-representable if it is representable over specific field $\mathbb{F}$.
\end{defn}

We will now define another important class of matroids. Due to its simplicity it will be suitable as an example in many of the notions defined in the proceeding text.

\begin{defn}\label{uniformM}
    Let $E = \{1, 2, \cdots, n\}$ and $\mathcal{I} = \{ L \subseteq E \; \text{ such that } \; |L| \leq m\}$. Then $(E, \mathcal{I})$ is a matroid which we denote by $U_{m,n}$ and call it a uniform matroid of rank $m$ on a $n$ element set.
\end{defn}
\begin{theorem}
    The ordered pair $U_{m,n} = (E, \mathcal{I})$ is a matroid
    
\end{theorem}

\begin{proof}
    
It is easy to check that $U_{m,n}$ is indeed a matroid. 


\begin{enumerate}
   

\item For any $m \geq 0$ we have $\emptyset \in \mathcal{I}$ since $|\emptyset | = 0$.

\item If $I \in \mathcal{I}$ and $J \subseteq I$ then $|J|\leq |I|$ so $|J|\leq |I| \leq m$ implying $J \in \mathcal{I}$ by definition.

\item  Finally, if $I, J \in \mathcal{I}$ and $|J|<|I|$ then for any $e \in I - J$ we will have $|J \cup e| = |J| + 1 \leq |I| \leq m$ so $J \cup e \in \mathcal{I}$ by definition for any $e \in I - J.$


\end{enumerate}
\end{proof}

\subsection{Bases}

Let $M = (E, \mathcal{I})$ be a matroid. We call a subset $B \subseteq E$ a \textit{basis} if it is a \textit{maximal independent set}. That means that $B$ is an independent set and $B$ is not properly contained inside any other independent set. It turns out that bases for matroids and bases for vector subspaces have some similarities in their properties. In particular the third property of independent sets immedietly guarantees us that all matroid bases have the same size, just like vector bases for finite-dimensional vector spaces.

\begin{theorem}
    Let $M = (E, \mathcal{I})$ be a matroid. Then all bases of $M$ have the same size.
\end{theorem}

\begin{proof}
    Aiming for contradiction, suppose that the theorem does not hold. Let $B$ and $S$ be two bases with $|B| < |S|$ - different size. By the third property of independent sets we know there exists $e \in S - B$ such that $ B \cup e \in \mathcal{I}$. However, then $B$ is properly contained inside $B \cup e$, another independent set, which contradicts its maximality. So the initial assumption that there exist two bases with different size is false.
\end{proof}

The concept of a basis is important because it allows us to define a \textit{rank function} of a matroid which is the size of the maximal independent subset inside a given subset of a matroid.  Because the sizes of all of such sets - bases - are equal, this will be a well-defined notion.

In figure \ref{1234bases} you find a depiction of the matroid we have seen in section 2.1. This time, we have added the bases of the matroid, colored red. Two properties can be observed. First of all, the bases are the maximal independent, which means that all of the sets they are strictly contained in are dependent. Second, both bases have two elements - the same size.

\begin{figure}[h]

\begin{center}
\begin{tikzpicture}

\matrix (a) [matrix of math nodes, column sep=0.6cm, row sep=0.6cm,]{
 & & &\textcolor{cyan}{
1234} & & & &\\
 \textcolor{cyan}{
123}& &\textcolor{cyan}{
124} & &\textcolor{cyan}{
134} &  & \textcolor{cyan}{
234}  \\
\textcolor{red}{12} & \textcolor{cyan}{13} & \textcolor{cyan}{14} & & \textcolor{red}{23} & \textcolor{cyan}{
24} & \textcolor{cyan}{
34} \\
\textcolor{orange}{1}& &\textcolor{orange}{2} & & \textcolor{orange}{3}& & \textcolor{cyan}{4} \\
& & & \textcolor{orange}{\emptyset} &  & & \\
&&&&&& \\};

\foreach \i/\j in {1-4/2-1, 1-4/2-3, 1-4/2-5, 1-4/2-7, 2-1/3-1, 2-1/3-2, 2-1/3-5, 2-3/3-1, 2-3/3-3, 2-3/3-6, 2-5/3-2, 2-5/3-3, 2-5/3-7, 2-7/3-5, 2-7/3-6, 2-7/3-7, 3-1/4-1, 3-1/4-3, 3-2/4-1, 3-2/4-5, 3-3/4-1, 3-3/4-7, 3-5/4-3, 3-5/4-5, 3-6/4-3, 3-6/4-7, 3-7/4-7, 3-7/4-5, 4-1/5-4, 4-3/5-4, 4-5/5-4, 4-7/5-4}
\draw[double, line width = 0.005mm, color = brown] (a-\i) -- (a-\j);

\node[draw] at (0, -2.5){\small \textcolor{orange}{Independent set}, \textcolor{red}{Bases} \textcolor{cyan}{Dependent set}};

\end{tikzpicture}
\end{center}
\caption{Diagram of a 4-element matroid}
\label{1234bases}
\end{figure}

Similar to the set of independent sets, we can characterize the set of bases of $M$ using the following two properties.

\begin{theorem}\label{basesproperties}
    Let $M$ be a matroid with $E$ as its ground set, and $\mathcal{B}$ be the set of bases of $M$, which we defined as the maximal independent sets. Then $\mathcal{B}$ satisfies the following properties.
    \begin{enumerate}
        \item[(B1)] $\mathcal{B}$ is non-empty
        \item[(B2)] If $B_1,B_2\in\mathcal{B}$ and $x\in B_1 - B_2$, then there exists $y\in B_2 - B_1$ such that $(B_1 - x)\cup y \in\mathcal{B}$.
    \end{enumerate}
\end{theorem}

Let us prove that the set of bases of a matroid satisfy the properties (B1) and (B2):
\begin{proof}
    To prove the first property observe that by defintion $\emptyset$ is always in $\mathcal{I}$ so the collection of independent sets is nonempty. This directly implies that the collection of bases is nonempty, because if we pick any independent set, there are finitely many independent sets it is included in. Than the we can pick the ones with the largest number of elements, since they are maximal. % for instance, an independent set with largest number of elements (there is at least one independent set! so we can talk about the largest one satisfying some properties) will be a basis, since it cannot be contained inside other independent set.
    
    
    For the second property let $x\in B_2 - B_1$. We first note that (I2) implies that $B_1 - x$ is independent, since it is a subset of $B_1$. Since $|B_1-x|<|B_2|$, we can apply (I3) on independent sets $B_1-x$ and $B_2$ to get that there exists a $y\in B_2-(B_1-x)$ such that $(B_1-x)\cup y \in\mathcal{I}$. Since $|(B_1-x)\cup y|=|B_1|$ and that all maximal independent sets have the same cardinality, $(B_1-x)\cup y$ must be a maximal independent set - thus a basis. Because $y$ is also in $B_2-B_1$, the property is fulfilled.
\end{proof} 



\begin{lemma}\label{samesize}
    Let $E$ be a finite set, and let $\mathcal{B}$ be a colection of subsets of $E$ satisfying (B1) and (B2). Then $\mathcal{B}$ is equicardinal, so all elements in $\mathcal{B}$ have the same cardinality.
\end{lemma}

Even though we have already proven that the set of bases of a matroid is equicardinal, we will also show that this property holds for an \textit{arbitrary collection of subsets} satisfying (B1) and (B2). We do this because we will need it in a future proof.

%Now that we have a new way to describe $\mathcal{B}$ with (B1) and (B2), we again need to prove that it is equicardinal (all elements of $\mathcal{B}$ having the same cardinality). 

\begin{proof}
    Suppose that $\mathcal{B}$ is not equicardinal, that is, we have $|B_1|<|B_2|$ for some $B_1,B_2\in\mathcal{B}$. From all $B_1,B_2$ where that holds, choose the pair that minimizes $|B_1-B_2|$. Since $B_1$ is bigger, choose an element $b\in B_1-B_2$ and apply (B2) to see that there exists a $d\in B_2-B_1$ such that $(B_1-b)\cup d \in\mathcal{B}$. With our choices of $b$ and $d$, we can see that $|((B_1-b)\cup d)-B_2|=|(B_1-b)-B_2|<|B_1-B_2|$. This, combined with that $|(B_1-b)\cup d|=|B_1|<|B_2|$, tells us that $B_1$ and $B_2$ are actually not the minimal choice. This contradiction proves that $\mathcal{B}$ is equicardinal.
\end{proof}

We are going to show the first cryptomorphism, referring to the idea that matroids can be defined not by the collection of independent sets but by its collection of bases. To prove that the two properties are sufficient to describe the bases of $M$, we will see that if $\mathcal{B}$ is any collection of subsets satisfying the bases axioms, then the independent sets should be all of the possible subsets of the the elements of $\mathcal{B}$.

\begin{theorem}
    Let $\mathcal{B}$ be a collection of subsets of a non-empty finite set $E$, satisfying (B1) and (B2). Let $\mathcal{I}=\{ I\subseteq B \; | \; B\in\mathcal{B} \}$. Then $(E,\mathcal{I})$ is a matroid with $\mathcal{B}$ as its set of bases.
\end{theorem}


\begin{proof}
    Our goal is of course to show that $\mathcal{I}$ satisfies the conditions of independent sets. If we manage to do that we will have that $(E, \mathcal{I})$ is a matroid having $\mathcal{B}$ as its collection of bases.
    \begin{enumerate}
        \item Since (B1) tells us that $\mathcal{B}$ is nonempty and because $\emptyset$ is a subset of any set, in particular $\emptyset \subseteq B$ where $B$ is some member of $\mathcal{B}$, we by definition have $\emptyset \in \mathcal{I}$.
        
        \item If $I\in\mathcal{I}$ and $J\subseteq I$, then by definition of $\mathcal{I}$, we have that $J\subseteq I\subseteq B$ for some $B\in\mathcal{B}$. Then we see that $J\subseteq B$ and by definition  $J\in\mathcal{I}$. So $\mathcal{I}$ satisfies (I2).
        
        \item For contradiction, suppose (I3) fails for some $I,J\in\mathcal{I}$ with $|I|<|J|$. This means that for all elements $x\in J-I$, $I\cup x \notin \mathcal{I}$. The idea of the proof will be to get $|I|\geq|J|$ for a contradiction. For this, we will have to derive a lot of ineqalities between cardinalities of the sets involved.
        
        Let $B_I$ and $B_J$ be the elements of $\mathcal{B}$ that contain $I$ and $J$, respectively, by definition of $\mathcal{I}$, we know such sets exist. Choose a $B_J$ so that $|B_J - (J\cup B_I)|$ is minimal. 
    
        With our choice of $I$ and $J$, it turns out that $J-B_I = J-I$. 
        The inclusion $J- B_I \subseteq J - I$ is clear because $I \subseteq B_I$. Therefore if $J-B_I \neq J-I$, there would have to be an element $e$ in $(J-I)-(J-B_I)$. We verify directly that this implies $e \in B_I$. Since $e$ is in $J-I$, our assumption that we made at the start tells us that $I\cup e\notin \mathcal{I}$, which contradicts $I\cup e\subseteq B_I$, since all subsets of a base are independent.

        Suppose now that $B_J-(J\cup B_I)$ is non-empty. Let $j$ be an element from this set, then (B2) tells us that there is an element $i\in B_I-B_J$ such that $(B_J-j)\cup i \in\mathcal{B}$. However, then $|((B_J-j)\cup i)-(J\cup B_I)|=|(B_J-j)-(J\cup B_I)|<|(B_J-(J\cup B_I)|$. We now found a new element $B=(B_J-j)\cup i$ in $\mathcal{B}$ that has $|B-(J\cup B_I)|$ smaller than $|B_J-(J\cup B_I|$, which contradicts that $B_J$ is the choice that minimalizes $|B_J-(J\cup B_I)|$. Thus, $B_J-(J\cup B_I)=\emptyset$. This now implies that $B_J\subseteq J\cup B_I$, which implies that $B_J-B_I\subseteq (J\cup B_I)-B_I=J-B_I$. Since $J-B_I\subseteq B_J-B_I$, this proves equality between $B_J-B_I$ and $J-B_I$.

        Next up we do the same thing, but this time proving that $B_I-(I\cup B_J)$ is empty. Assuming the opposite by letting $i$ be a part the set, (B2) tells us that there exists a $j\in B_J-B_I=J-B_I=J-I$ such that $(B_I-i)\cup j\in\mathcal{B}$. Since our choice of $i$ tell us that $i \notin I$ we have that $I\cup j\subseteq (B_I-i)\cup j$, it means $I\cup j\in\mathcal{I}$. This gives us precicely (I3), which contradicts our originil assumption. Thus, $B_I-(I\cup B_J)=\emptyset$. By the same logic we used in the previous paragraph, $B_I-B_J=I-B_J\subseteq I-J$.

        We can use the fact that $\mathcal{B}$ is equicardical, that is \ref{samesize}, to see that $|B_I-B_J|=|B_J-B_I|$. At last, we use everything obtain to get the following
        $$ |I-J|\geq |B_I-B_J|=|B_J-B_I|=|J-B_I|=|J-I|. $$

        Therefore we obtain the inequality $|I - J|\geq |J - I|.$

        Because we know that $ I = (I - J) \cup (I\cap J)$ and clearly $I - J$ and $I \cup J$ are disjoint we see that

        $$ |I| = |(I - J)\cup (I \cap J)| = |I - J| + |I \cap J| - |\underbrace{(I - J) \cap (I \cap J)}_{ = \emptyset}|.$$

        Therefore $|I| = |I-J| + |I \cap J|$ and by the same reasoning $|J| = |J - I| + |I\cup J|$. Now we are almost done, combining the above ineqalities with $|I - J|\geq |J - I|$ we observe that $|I|\geq |J|$. This contradicts our assumption that $|I|< |J|$. 
        
        Thus we have derived a contradiction when assuming that there exits, a pair $I, J$ of subsets violating (I3). Hence$\mathcal{I}$ satisfies (I3) and all in all $(E,\mathcal{I})$ is a matroid.

        
    \end{enumerate}  

    
    The fact that $\mathcal{B}$ is now the set of bases of the matroid $(E,\mathcal{I})$ follows from the fact that all elements in $\mathcal{B}$ are maximal in $\mathcal{I}$.
\end{proof}



\newpage

\subsection{Circuits} 
As mentioned before we can define a matroid in different forms, one of this other definitions is in terms of circuits. Before the definition, we need to introduce an additional concept. 
That is, a minimal dependent set, these are dependent sets whose all proper subsets are independent.

Now, we can state the following:
\begin{defn}
Let $M = (E, \mathcal{I})$ be a matroid. Any subset $D \in \mathcal{I}$ which is not independent is called dependent. Circuit will be a minimal dependent set of an arbitrary matroid $M$. The collection will be denoted by $C$ or $\mathcal{C}(M)$. Additionally, we define a circuit of a matroid $M$, that has $n$ elements, as an $n$-circuit.
\end{defn}

The name is a reference to matroid circuits corresponding to circuits of the underlying graph when talking about graphic matroids.

Similarly, as with the independent sets, we can use the circuits to formulate a definition for a given matroid. And moreover, in the same way we can use the independent sets of a matroid to determine its circuits, we can use the circuits of a matroid to determine its independent sets.

So, we can characterize the concept of a matroid in the following way:

\begin{defn}
    Let $E$ be a non-empty finite set, and $\mathcal{C}$ a collection of subsets of $E$, called circuits, such that:

    \begin{enumerate}
        \item[C1)] $\emptyset \notin \mathcal{C}$

        \item[(C2)] If $C_1$ and $C_2$ are members of $\mathcal{C}$ and $C_1 \subseteq C_2$, then $C_1 = C_2$.

        \item[(C3)] If $C_1$ and $C_2$ are distinct members of $\mathcal{C}$ and there exist an $e \in C_1 \cap C_2$, then there is a member $C_3$ of $\mathcal{C}$, such that $C_3 \subseteq (C_1  \cup C_2) - e$
    \end{enumerate}
    
\end{defn}

Again, we will use the same matroid example as before, this time adding circuits to the drawing (in blue). As you can see, the circuits are minimal dependent sets where everything above it is a dependent set. Also note how all independent sets do not contain any circuit.

\begin{center}
\begin{tikzpicture}

\matrix (a) [matrix of math nodes, column sep=0.6cm, row sep=0.6cm,]{
 & & &\textcolor{cyan}{
1234} & & & &\\
 \textcolor{cyan}{
123}& &\textcolor{cyan}{
124} & &\textcolor{cyan}{
134} &  & \textcolor{cyan}{
234}  \\
\textcolor{red}{12} & \textcolor{blue}{13} & \textcolor{cyan}{14} & & \textcolor{red}{23} & \textcolor{cyan}{
24} & \textcolor{cyan}{
34} \\
\textcolor{orange}{1}& &\textcolor{orange}{2} & & \textcolor{orange}{3}& & \textcolor{blue}{4} \\
& & & \textcolor{orange}{\emptyset} &  & & \\
&&&&&& \\};

\foreach \i/\j in {1-4/2-1, 1-4/2-3, 1-4/2-5, 1-4/2-7, 2-1/3-1, 2-1/3-2, 2-1/3-5, 2-3/3-1, 2-3/3-3, 2-3/3-6, 2-5/3-2, 2-5/3-3, 2-5/3-7, 2-7/3-5, 2-7/3-6, 2-7/3-7, 3-1/4-1, 3-1/4-3, 3-2/4-1, 3-2/4-5, 3-3/4-1, 3-3/4-7, 3-5/4-3, 3-5/4-5, 3-6/4-3, 3-6/4-7, 3-7/4-7, 3-7/4-5, 4-1/5-4, 4-3/5-4, 4-5/5-4, 4-7/5-4}
\draw[double, line width = 0.005mm, color = brown] (a-\i) -- (a-\j);

\node[draw] at (0, -2.5){\small \textcolor{orange}{Independent set}, \textcolor{red}{Basis}, \textcolor{cyan}{Dependent set}, \textcolor{blue}{Circuit}};
% \caption{Figure 3: Diagram of a 4-element matroid}
\end{tikzpicture}
\end{center}

To get an intuition behind (C3), let us consider the matroid induced by the graph in figure 3. We will discuss graphic matroid later, but just know that a circuit is a cycle in the graph. Here, we have two circuits, $C_1$ and $C_2$, that share an edge, $e$. If you take the union of both circuits and remove the edge $e$ from it, then that subgraph contains another cycle, circuit $C_3$.

% Someone please fix this!


\begin{figure}
\caption{A graph of suspicious shape, having three cycles.}
\begin{subfigure}[h]{0.245\textwidth}
  \caption{Graph}
\begin{tikzpicture}[node distance={10mm}, thick, main/.style = {draw,circle,fill,inner sep=1pt}] 
  \node[main] (1) at (0,0) {}; 
  \node[main] (3) at (-0.5,0.5) {}; 
  \node[main] (4) at (0, 1) {}; 
  \node[main] (5) at (0, -1) {}; 
  \node[main] (6) at (1.5, -1) {}; 
  \node[main] (7) at (0.75, 1.5) {}; 
  \node[main] (8) at (1.5, 1) {}; 
  \node[main] (9) at (0.5, -1) {}; 
  \node[main] (10) at (1, -1) {}; 
  \node[main] (11) at (0.75, -0.5) {}; 
  \draw (1) to [bend left] (3);
  \draw (1) to [bend right=90]
node[midway, right] {e} (4) ;
  \draw (3) to [bend left] (4);
  \draw (4) to [bend left] (7);
  \draw (7) to [bend left] (8);
  \draw (8) to (6);
  \draw (5) to (9);
  \draw (6) to (10);
  \draw (9) to [bend left] (11);
  \draw (10) to [bend right] (11);
  \draw (5) to (1);
\end{tikzpicture} 
\end{subfigure}
\begin{subfigure}[h]{0.245\textwidth}
  \caption{$C _1$}
\begin{tikzpicture}[node distance={10mm}, thick, main/.style = {draw,circle,fill,inner sep=1pt}] 
  \node[main] (1) at (0,0) {}; 
  \node[main] (3) at (-0.5,0.5) {}; 
  \node[main] (4) at (0, 1) {}; 
  \node[main] (5) at (0, -1) {}; 
  \node[main] (6) at (1.5, -1) {}; 
  \node[main] (7) at (0.75, 1.5) {}; 
  \node[main] (8) at (1.5, 1) {}; 
  \node[main] (9) at (0.5, -1) {}; 
  \node[main] (10) at (1, -1) {}; 
  \node[main] (11) at (0.75, -0.5) {}; 
  \draw (1) to [bend right=90]
node[midway, right] {e} (4) ;
  \draw (4) to [bend left] (7);
  \draw (7) to [bend left] (8);
  \draw (8) to (6);
  \draw (5) to (9);
  \draw (6) to (10);
  \draw (9) to [bend left] (11);
  \draw (10) to [bend right] (11);
  \draw (5) to (1);
\end{tikzpicture} 
\end{subfigure}
\begin{subfigure}[h]{0.245\textwidth}
  \caption{$C _2$}
\begin{tikzpicture}[node distance={10mm}, thick, main/.style = {draw,circle,fill,inner sep=1pt}] 
  \node[main] (1) at (0,0) {}; 
  \node[main] (3) at (-0.5,0.5) {}; 
  \node[main] (4) at (0, 1) {}; 
  \node[main] (5) at (0, -1) {}; 
  \node[main] (6) at (1.5, -1) {}; 
  \node[main] (7) at (0.75, 1.5) {}; 
  \node[main] (8) at (1.5, 1) {}; 
  \node[main] (9) at (0.5, -1) {}; 
  \node[main] (10) at (1, -1) {}; 
  \node[main] (11) at (0.75, -0.5) {}; 
  \draw (1) to [bend left] (3);
  \draw (1) to [bend right=90]
node[midway, right] {e} (4) ;
  \draw (3) to [bend left] (4);
\end{tikzpicture} 
\end{subfigure}
\begin{subfigure}[h]{0.245\textwidth}
  \caption{$C _3$} 
\begin{tikzpicture}[node distance={10mm}, thick, main/.style = {draw,circle,fill,inner sep=1pt}] 
  \node[main] (1) at (0,0) {}; 
  \node[main] (3) at (-0.5,0.5) {}; 
  \node[main] (4) at (0, 1) {}; 
  \node[main] (5) at (0, -1) {}; 
  \node[main] (6) at (1.5, -1) {}; 
  \node[main] (7) at (0.75, 1.5) {}; 
  \node[main] (8) at (1.5, 1) {}; 
  \node[main] (9) at (0.5, -1) {}; 
  \node[main] (10) at (1, -1) {}; 
  \node[main] (11) at (0.75, -0.5) {}; 
  \draw (1) to [bend left] (3);
  \draw (3) to [bend left] (4);
  \draw (4) to [bend left] (7);
  \draw (7) to [bend left] (8);
  \draw (8) to (6);
  \draw (5) to (9);
  \draw (6) to (10);
  \draw (9) to [bend left] (11);
  \draw (10) to [bend right] (11);
  \draw (5) to (1);
\end{tikzpicture} 
\end{subfigure}
                  \end{figure}

We can say that a set $X \subseteq E$ is independent if and only if, it does not contain any circuit.

The concept of circuits points towards the direction of something similar to a complement of the independent sets, but not completely, since it is only the minimal dependent, and not all dependet subsets. We now have the following theorem.

\begin{lemma}\label{lem:circuits-have-properties}
  All circuits in a matroid satisfy (C1), (C2) and (C3).
\end{lemma}

\begin{proof} We will prove each property individually:
  
\begin{enumerate}
    \item $\emptyset \in \mathcal I$, therefore $\emptyset$ is not dependent (which means it cannot be a circuit)
    \item A counterexample to this clause would violate the $C _2$ being minimal.
    \item Assume a counterexample exists. Because $C _1 \cup C _2 - e$ does not contain a circuit, it must be an element of $\mathcal I$. 

        We claim that $C _2 \setminus C _1$ is nonempty. Assume this is not the case. That would imply $C _2 \subseteq C _1$, but $C _1 \neq C _2 $ by (C3), which means that $C _1 $ is a proper subset of $C _2$. As both $C _1 $ and $C _2 $ are members of $\mathcal C$, this violates (C2), and is a contradiction. We can then assume some $f \in C _2 \setminus C _1 $ exists.

        As $C _2$ is minimally dependent, we know that $C _2 - f$ is independent. Let $I$ be a maximal independent subset of $C _2 \cup C _1$ which is a superset of $C _2 - f$. Clearly, $f \not\in I$ (otherwise $C _2 \subseteq I \in \mathcal I$). Furthermore, as $C _1 $ is a circuit, we know that $C _1 \not \subseteq I \in \mathcal I$ (otherwise $C _1$ would be independent). This means some $g \in C _1$ exists such that $g \not\in I$. As $f \in C _2  \setminus  C _2$, we know that $f \neq g$.

        As $I \subseteq C _1 \cup C _2$ and $f, g \in C _1 \cup C _2$ but $f, g \not\in I$, we can infer that 
        \begin{align*}
        |I| \leq |C _1 \cup C _2 - f - g| = |C _1 \cup C _2| - 2.
        \end{align*}

        We also notice that $|C _1 \cup C _2 - e| = |C _1 \cup C _2 | - 1$. We can now conclude by transitivity with $|C _1 \cup C _2 | - 2 < |C _1 \cup C _2 | - 1$ that $|I| < |C _1 \cup C _2 - e|$. We recall that both $I$ and $C _1 \cup C _2 - e$ are members of $\mathcal I$. We can apply (I3) to generate a superset of $I$ which contradicts it's maximallity.
\end{enumerate}
\end{proof}

\begin{theorem}\label{thm:matroid-circuit-definition}
Let $E$ be a set and $\mathcal C$ be a collection of subsets of $E$ which satisfies the conditions outlined above. Let  $\mathcal I$  be the collection of subsets of $E$ that contain no member of $\mathcal C$, that is 

\begin{align}
   % \forall X \in \mathcal{I}, \forall C \in \mathcal C,  C \not \subseteq X. 
   \mathcal{I} = \{I \in 2^E |\; \text{for all } \; C \in \mathcal{C}\; \text{we have} \; C \not\subseteq I\}
    \label{independent-sets-from-circuits}
\end{align}

    The pair $(E,\mathcal I)$ is a matroid having $\mathcal C$ as its collection of circuits.
\end{theorem}

\begin{proof} We start by showing that $\mathcal I $ satisfies the necessary conditions for $(E, \mathcal I )$ to be a matroid:
    \begin{enumerate} 
        \item The only subset of $\emptyset $ is $ \emptyset $, but $ \emptyset \not\in \mathcal C $, so $ \emptyset $ satisfies (\ref{independent-sets-from-circuits}), hence $\emptyset \in \mathcal I$.
    \item Assume we have $X \subseteq Y \in \mathcal I$. Assume that $X \not\in I$, i.e.\ there exists some $C \in \mathcal C $ such that $C \subseteq X$. We recall that $\subseteq $ is transitive, thus $C \subseteq Y$ and $Y \not\in I$, hence a contradiction.
    \item Given $X, Y \in \mathcal I $ with $|X| < |Y|$ we must show there exists some $e \in Y \setminus X$ such that $X \cup \{e\} \in \mathcal I$. Assume that such an $e$ does not exist, i.e.\ for all $e \in Y \setminus X$ we have some circuit $C _e \in \mathcal C $ such that $C _e \subseteq X \cup \{e\}$. We obviously have $e \in C_e$, because otherwise $C_e \subseteq X$ and $X \not\in \mathcal I$.


     Let $X' = X \setminus Y$ and $Y' = Y \setminus X$. We will prove the statement by induction: given some set $S \subseteq X'$, a set of circuits $D \subseteq \mathcal C$ with $\forall C \in D, C \subseteq  S \cup Y$ and a surjective but not injective function $f : D \to S$ such that $f(D)$ is maximal in size and $\forall C \in D, f(C) \in C$, our statement is proven. We will use induction on the size of $S$:
    \begin{enumerate}
      \item If $S = \{e\}$ only has one element, we know that $f$ is not injective, so we must have distinct $C _1, C _2 \in D$ such that $e = f(C _1 ) = f(C _2)$. We can use the $3$-rd circuit axiom to generate some circuit $C _3 \subseteq C _1 \cup C _2 - e$. We recall that $C _1 , C _2 \subseteq S \cup Y $, so $C _3 \subseteq S \cup Y - e$ which implies $C _3 \subseteq Y$, which is a contradiction.
        \item Inductive step. Assuming the statement is true for all choices of $S$ with $k$ elements, we will attempt to prove it is also true when $S$ has $k + 1$ elements. Recalling that $f$ is not injective, we must have distinct $C _1, C _2 \in D$ such that $f(C _1 ) = f(C _2)$. Let $g = f(C _1)$. We have $g \in C _1, C _2$, so by the $3$-rd circuit axiom we must have some circuit $C  _3 \subseteq C _1 \cup C _2 - g$. Let $S' = C _3 \cap X'$. 

            For all $e \in S'$, we claim there must exist some $C_e \in D$ such that $e = f(C_e)$. Assume this is not the case, i.e.\ there is some $e \in S'$ such that no valid choice of $C_e$ exists. Given that $e \in S'$, we know that either $e \in C_1$ or $e \in C_2$. After a potential swap, we assume that $e \in C_1$. We can now define $f'(C) = e$ when $C = C_1$ and $f'(C) = f(C)$. Because no choice of $C_e$ exists, we know that $e \not\in f(D)$. Because $f(C _1 ) = f(C _2 )$ we have $f(D) = f(D - C _1 )$ which lets us compute $f'(D) = f(D) \cup \{e\}$, which means $f(D)$ was not maximal, hence a contradiction. Our claim is then true.

            Define $D' = \{C_e | e \in S'\} \cup \{C _3 \}$. We've essentially proven in the last paragraph that $f : \{C_e | e \in S'\} \to S'$ is still a  surjective function. It then follows that $h = f : D' \to S'$ is surjective as well. By the Pigeonhole principle, we know $h$ cannot be injective, because $|D'| = |S'| + 1$. We are now almost ready to apply the induction hypothesis to $(S', D', h)$. We've shown a choice for $h$ exist. If this choice is not maximal, swap it with a maximal choice (we had to show that at least one choice exists for a maximal choice to exist). We can now apply the induction hypothesis, proving the statement.
    \end{enumerate}

    We notice that $X$ can be partitioned into $X'$ and $X \cap Y$, hence $|X| = |X'| + |X \cap Y|$. We can follow a similar process for $Y$. Combining the two together with the fact that $|X| < |Y|$ yields that $|X'| < |Y'|$. We know that $Y \in \mathcal I$ so $\forall e \in Y', C_e \not \subseteq  Y$, but $C_e \subseteq  X \cup \{e\}$ therefore there must exist some $c_e \in C_e \cap X'$. 

    We define $D = \{C_e, e \in Y'\}$, $f(C_e) = c_e$ and $S = f(D)$. Furthermore, we know that $f$ is not injective by the Pigeonhole principe (we have $|D| = |Y'| > |X'| \geq |S|$). We also know, by the definition of $S$, that $f$ is surjective. We've shown a choice of $f$ exists. If it is not maximal, swap it with a maximal choice. We can now apply the statement proven by induction to finish the proof.
\end{enumerate}


We now show that $C$ is a circuit for the newly defined matroid if and only if it is a member of $\mathcal C$. 
\begin{enumerate}
  \item[$\implies$]
  If $C$ is a circuit, then $C \not\in \mathcal I$, which means that $C$ contains some $D \in \mathcal C$. As $C$ is minimally dependent, all it's proper subsets are independent, but members of $\mathcal C$ cannot be independent (as seen in the definition of $\mathcal I$), which implies that $C = D$, hence $C \in \mathcal C$.
  \item[$\impliedby$]
  Let $C \in \mathcal C$. Assume $C$ is not a circuit. Clearly, $C \not\in \mathcal I$ is implied by the definition of $\mathcal I$ seen above. This implies that $C$ is dependent but not a circuit, which means $C$ is not miminal. 

  A proper subset $D  \subseteqneq C$ must then exist such that $D$ is a circuit. We've already shown in the $\implies$ direction how this implies $D \in \mathcal C$. We have now constructed a contradiction to (C2), hence our assumption that such a circuit exists is false.

\end{enumerate}



  We can now conclude that $(E, \mathcal I)$ is indeed a matroid with $\mathcal C$ as it's set of circuits.
\end{proof}

\input{graphs}

\section{Rank Function}

The rank function is a function $r:2^E \rightarrow \mathbb{Z}_{\geq0}$ that, when given a $X\subseteq E(M)$, gives you the cardinality of the maximal independent set contained in $X$. In other words:
$$ r(X) = \max\{\, |I| : I\in\mathcal{I} \text{ and } I\subseteq X \} $$
For example, $r(M):=r(E(M))=|B|$ for some $B\in\mathcal{B}(M)$. Like all the previous times, we can characterize the rank function with the following properties:

\begin{defn}
    Let $E$ be a non-empty finite set, and let $\rank : 2^E \rightarrow \mathbb{Z}_{\geq0}$ be a function satisfying:
    \begin{enumerate}
        \item[(R1)] If $X\subseteq E$, then $0 \leq \rank(X) \leq |X| $
        \item[(R2)] If $X\subseteq Y\subseteq E$, then $\rank(X)\leq\rank(Y)$
        \item[(R3)] If $X,Y\subseteq E$, then $\rank(X\cup Y)+\rank(X\cap Y) \leq \rank(X)+\rank(Y) $
    \end{enumerate}
\end{defn}

Below, we again find our example matroid drawing. This time, instead of using a new color, we will denote the rank of each set by writing it in subscript below the set. We can see that the rank of any set is atmost its cardinality and each subset of some set has at most the rank of that set. As always it takes a bit more effort to show that the last property holds for a given matroid.

\begin{center}
\begin{tikzpicture}

\matrix (a) [matrix of math nodes, column sep=0.6cm, row sep=0.6cm,]{
 & & &\textcolor{cyan}{
1234}_2 & & & &\\
 \textcolor{cyan}{
123}_2& &\textcolor{cyan}{
124}_2 & &\textcolor{cyan}{
134}_1 &  & \textcolor{cyan}{
234}_2  \\
\textcolor{red}{12}_2 & \textcolor{blue}{13}_1 & \textcolor{cyan}{14}_1 & & \textcolor{red}{23}_2 & \textcolor{cyan}{
24}_1 & \textcolor{cyan}{
34}_1 \\
\textcolor{orange}{1}_1& &\textcolor{orange}{2}_1 & & \textcolor{orange}{3}_1& & \textcolor{blue}{4}_0 \\
& & & \textcolor{orange}{\emptyset}_0 &  & & \\
&&&&&& \\};

\foreach \i/\j in {1-4/2-1, 1-4/2-3, 1-4/2-5, 1-4/2-7, 2-1/3-1, 2-1/3-2, 2-1/3-5, 2-3/3-1, 2-3/3-3, 2-3/3-6, 2-5/3-2, 2-5/3-3, 2-5/3-7, 2-7/3-5, 2-7/3-6, 2-7/3-7, 3-1/4-1, 3-1/4-3, 3-2/4-1, 3-2/4-5, 3-3/4-1, 3-3/4-7, 3-5/4-3, 3-5/4-5, 3-6/4-3, 3-6/4-7, 3-7/4-7, 3-7/4-5, 4-1/5-4, 4-3/5-4, 4-5/5-4, 4-7/5-4}
\draw[double, line width = 0.005mm, color = brown] (a-\i) -- (a-\j);

\node[draw] at (0, -2.5){\small \textcolor{orange}{Independent set}, \textcolor{red}{Basis}, \textcolor{cyan}{Dependent set}, \textcolor{blue}{Circuit}, $\text{Set}_{\rank(\text{Set})}$ };

\end{tikzpicture}
\end{center}

Now, let us prove that the rank function of a matroid satirfy these properties:

\begin{proof}
    \,
    \begin{enumerate}
        \item Since the co-domain is $\mathbb{Z}_{\geq0}$, $0\leq\rank(X)$. Since the maximal independent set contained in $X$ is... contained in $X$, the cardinality of that set must be smaller or equal to $|X|$. Thus, $0\leq\rank(X)\leq|X|$.
        \item Let $I_X$ be a minimal independent set contained in $X$, so $\rank(X)=|I_X|$. $I_X\subseteq X\subseteq Y$ implies that $I_X$ is an independent set contained in $Y$. Since $I_X$ may or may not be a maximal independent set contained in $Y$, $\rank(X)=|I_X|\leq\rank(Y)$. 
        \item We assume that $A\subseteq B\subseteq E$. Let $I_{A\cap B}$ be a maximal independent set contained in $A\cap B$. Since $A\cap B\subseteq A\cup B$, $I_{A\cap B}$ must be an independent set contained in $A\cup B$. Let $I_{A\cup B}$ be $I_{A\cap B}$ extended to be a \textbf{maximal} independent set contained in $A\cup B$. 
        
        Now let us take a look at $I_{A\cup B}\cap A$. Since the set is contained in $A$, (I2) tells us that $I_{A\cup B}\cap A$ is in independent set contained in $A$. Since the set may or may not be maximal in that regard, $|I_{A\cup B}\cap A|\leq \rank(A)$. Similarly, $|I_{A\cup B}\cap B|\leq \rank(B) $.
        
        $$ \rank(X)+\rank(Y) \geq |I_{A\cup B}\cap A|+|I_{A\cup B}\cap B| $$
        $$ = |(I_{A\cup B}\cap A)\cup(I_{A\cup B}\cap B)|+|(I_{A\cup B}\cap A)\cap(I_{A\cup B}\cap B)| $$
        $$ = |I_{A\cup B}\cap (A\cup B)|+|I_{A\cup B}\cap (A\cap B)| $$

        Since $I_{A\cup B}\subseteq A\cup B$, $I_{A\cup B}\cap(A\cup B) = I_{A\cup B}$.

        Since $I_{A\cap B}\subseteq I_{A\cup B}$, $I_{A\cup B}\cap (A\cap B) \supset I_{A\cap B}\cap (A\cap B) = I_{A\cap B} $. Let $e\in I_{A\cup B}\cap (A\cap B)$. Aiming for contradiction, suppose $e\notin I_{A\cap B}$. Since $I_{A\cap B}\cup e \subseteq I_{A\cup B} $, $I_{A\cap B}\cup e$ must be an independent set conained in $A\cap B$. This contradicts the fact that $I_{A\cap B}$ is a maximal in that regard. Thus, $e\in I_{A\cap B}$. This proves that $I_{A\cup B}\cap(A\cap B)\subseteq I_{A\cap B} $, which implies equality between the two sets.

        Thus,

        $$ \rank(X)+\rank(Y) \geq |I_{A\cup B}|+|I_{A\cap B}| = \rank(A\cup B)+\rank(A\cap B) $$
    \end{enumerate}
\end{proof}
One might observe that the rank of an independent set is the cardinality of the set itself. This is because the largest independent set that is contained in this independent set is ofcourse the independent set itself. With this property, we can define the set of independent sets with the three properties of the rank function.
\begin{theorem}
    Let $E$ be a finite set, and let $\rank : 2^E \rightarrow \mathbb{Z}_{\geq0}$ be a function satisfying (R1)-(R3). Let $\mathcal{I} = \{ I\subseteq E : \rank(I)=|I| \} $. Then $(E,\mathcal{I})$ is a matroid with $\rank$ as its rank function.
\end{theorem}
Before we prove this, let's prove this small theorem first:
\begin{theorem}
\label{rankextension}
    Let $E$ be a finite set and $\rank:2^E\rightarrow \mathbb{Z}_{\geq0}$ be a function satisfying (R2) and (R3). Let $X,Y\subseteq E$. Then if for all $y\in Y-X$, $\rank(X\cup y)=\rank(X)$, then $\rank(X\cup Y)=\rank(Y)$.
\end{theorem}
\begin{proof}
    We will prove that the statement holds with $Y-X = \{a_1,a_2,\dots,a_k\}$ holds for all itegers $k$.
    \begin{enumerate}
        \item Base case: $k=1$

        If $\rank(X\cup a_1)=\rank(X)$, then $\rank(X\cup Y)=\rank(X\cup(Y-X))=\rank(X\cup a_1)=\rank(X)$.
        \item Induction Step: Assume the statement holds for $k=n$.
        \item Let $k=n+1$: Assume that $Y-X=\{a_1,\dots,a_{n+1}\}$ for all $i\in\{1,\dots,n+1\}$ $\rank(X\cup a_i)=\rank(X)$. Then, using the assumption and our induction assumption, 
        $$ \rank(X)+\rank(X) = \rank(X\cup \{a_1,\dots,a_n\})+\rank(X\cup a_{n+1}) $$
        Using (R3), we get
        $$ \geq \rank([X\cup\{a_1,\dots,a_n\}]\cup [X\cup a_{n+1}]) + \rank([X\cup\{a_1,\dots,a_n\}]\cap [X\cup a_{n+1}]) $$
        $$ = \rank(X\cup\{a_1,\dots,a_{n+1}\}) + \rank(X) $$
        Using (R2), we get
        $$ \geq \rank(X)+\rank(X) $$
        Since $\rank(X)+\rank(X)$ is on both sides, equality holds throughout. Thus, $\rank(X\cup\{a_1,\dots,a_{n+1}\})=\rank(X\cup Y)=\rank(X)$. And thus, the statement holds for $k=n+1$.
    \end{enumerate}
    Thus, by mathematical induction, the statement holds for all integers $k$.
\end{proof}
Finally, we can prove Theorem 5:
\begin{proof}
    To prove that $(E,\mathcal{I})$ is a matroid, we will check if $\mathcal{I}$ satisfies (I1)-(I3).
    \begin{enumerate}
        \item (R1) tells us that $0\leq\rank(\emptyset)\leq|\emptyset|=0$, which implies that $\rank(\emptyset)=0=|\emptyset|$, which means that $\emptyset\in\mathcal{I}$, so (I1) is satisfied.
        \item Assume we have $J\subseteq I\in\mathcal{I}$. Then, using (R3),
        $$ \rank(J)+\rank(I-J) \geq \rank(J\cup[I-J])+\rank(J\cap[I-J]) = \rank(I)+\rank(\emptyset)=|I| $$
        Using (R2), we also get that
        $$ \rank(J)+\rank(I-J) \leq |J|+\rank(I-J) \leq |J|+|I-J| = |I| $$
        This implies that equality must hold throughout. At last, we show that
        $$ \rank(J)+\rank(I-J) = |J|+\rank(I-J) \Rightarrow \rank(J)=|J| $$
        Thus, $J\in\mathcal{I}$, so {I2} is satisfied.
        
        \item Suppose (I3) does not hold for some $I,J\in\mathcal{I}$ with $|I|>|J|$. Then for all $e\in I-J$, $J\cup e\notin \mathcal{I}$. Also, remember that $I,J\in\mathcal{I}$ means that $\rank(I)=|I|$ and $\rank(J)=|J|$.

        Since $J\subseteq J\cup e$, we can use (R1) and (R2) to see that
        $$ |J|=\rank(J)\leq\rank(J\cup e)|\leq|J\cup e|=|J|+1 $$
        $J\cup e\notin \mathcal{I}$ tells us that $\rank(J\cup e)\neq|J\cup e|$, so that must mean that $\rank(J\cup e)$ is equal to what's on the left hand side, so $\rank(J)$.

        This gives us that for all $e\in I-J$, $\rank(J\cup e)=\rank(J)$, which, according to Theorem 6, implies that $\rank(J\cup I)=\rank(J)=|J|$. However, since $I\subseteq J\cup I$, (R2) tells us that $\rank(J\cup I) \geq \rank(I) = |I|$, which means that $|J|\geq|I|$. This contradicts our assumption that $|J|<|I|$. Thus, (I3) is satisfied.
    \end{enumerate}
    Thus, $(E,\mathcal{I})$ is a matroid.
\end{proof}

\begin{exmp}
    \input{U24}
\end{exmp}



